---
title: "C4TS R-Club - Session 2"
author: "Jason Pott"
date: "30/06/2019"
output: rmarkdown::github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# If you have not installed the tidyverse then enter "install.packackages("tidyverse") " in the console
library(tidyverse)


```

In our [first session](https://github.com/jasonpott/r4ds/blob/master/01.%20Session%201%20/session_1.md) we covered - Setting up our environment
- installing packages
- introduction to ggplot
- I gave an example of how to load a csv file into R

Feel free to look back over the notes from that session, please come back to me if you have any questions or need some clarification.


### Session 2
In this session we will work through some more useful ways of working in R and Rstudio. 

If you were not here last week the only prerequisite is to make your Rstudio look like mine. Use the menus: Menu > View > Panes > Pane Layout, to acomplish this.

This week we will discuss:
- Working in projects and why you should
- Getting data into R
- Manipulating data
- creating tables
- passing manipulated data to ggplot

## Projects in R and Rstudio

```{r, include=TRUE, echo=TRUE}

```


## Importing data into R

```{r, include=TRUE, echo=TRUE}

```
