---
title: "C4TS R-Club - Session 2"
author: "Jason Pott"
date: "30/06/2019"
output: rmarkdown::github_document
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```



In our [second session](https://github.com/jasonpott/r4ds/blob/master/02.%20Session%202/session_2.md) we covered - Setting up our environment
- installing packages
- introduction to ggplot
- I gave an example of how to load a csv file into R

Feel free to look back over the notes from that session, please come back to me if you have any questions or need some clarification.


## Session 3

This session will cover

- select()  
- mutate()  
- group_by()  
- tally()  
- transmute()  

We will also create some tables then pass some data to ggplot

select()

```{r, include=TRUE, echo=TRUE}

```

mutate()
```{r, include=TRUE, echo=TRUE}

```

group_by()
```{r, include=TRUE, echo=TRUE}

```

tally()
```{r, include=TRUE, echo=TRUE}

```


transmute()
```{r, include=TRUE, echo=TRUE}

```

### Creating tables



```{r, include=TRUE, echo=TRUE}
```

### passing manipulated data to ggplot


```{r, include=TRUE, echo=TRUE}
```